A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities

### 2011-07-15 papers/contact/paper.tex
add a few figures for contact density
add a bunch more contact-density visualization and new c.d. functional

@@

>>>>

\section{Comparison with simulation}

\subsection{Hard spheres confined in a spherical cavity}

\begin{figure}
  \includegraphics[height=5cm]{figs/sphere-radial-06-04}
  \caption{Density and contact densities of four hard spheres in a
    spherical cavity with radius 6.}
  \label{fig:sphere-density}
\end{figure}

\begin{figure}
  \includegraphics[height=5cm]{figs/sphere-radial-08-13}
  \caption{Density and contact densities of fourteen hard spheres in a
    spherical cavity with radius 8.}
  \label{fig:sphere-density2}
\end{figure}

\subsection{Hard spheres confined in a cubical cavity}

\begin{figure*}
\includegraphics[height=5cm]{figs/box-100c--05,05,05-13}
\includegraphics[height=5cm]{figs/box-100s--05,05,05-13}
\includegraphics[height=5cm]{figs/box-110c--05,05,05-13}
\caption{Density of hard spheres in a $5\times5\times5$ box.}
\label{fig:box-density}
\end{figure*}

@@

\section{Derivatives of the fundamental-measure weighted densities}

The derivative of the filling fraction is very easy:
\begin{align}
  n_3(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
  \Theta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
  \frac{\delta n_3(\mathbf{r})}{\delta R(\mathbf{r}')}
  &= n(\mathbf{r'})\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
\end{align}

The derivative of the density of surface is harder:
\begin{align}
  n_2(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
  \frac{\delta n_2(\mathbf{r})}{\delta R(\mathbf{r}')}
  &= n(\mathbf{r'})\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
\end{align}
where the trouble here is that the derivative of the delta function,
$\delta'$, isn't something we really like to deal with.  Particularly
as we're looking  at a 3D integral involving a 1D delta function,
which means there are Jacobian factors to be considered.  On the other
hand, there's no reason why we couldn't convolve with this kernel as
efficiently as we do with any other kernel.

The derivatives oof the remaining scalar densities can be reduced to
sums of the terms above:
\begin{align}
  n_1(\mathbf{r}) &= \int \mathbf{dr'} \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}
  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
  \frac{\delta n_1(\mathbf{r})}{\delta R(\mathbf{r}')}
  &= \frac{n(\mathbf{r'})}{4\pi
    R(\mathbf{r'})}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
  -
  \frac{n(\mathbf{r'})}{4\pi
    R(\mathbf{r'})^2}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
\end{align}
\begin{align}
  n_0(\mathbf{r}) &= \int \mathbf{dr'} \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})^2}
  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
  \frac{\delta n_0(\mathbf{r})}{\delta R(\mathbf{r}')}
  &= \frac{n(\mathbf{r'})}{4\pi
    R(\mathbf{r'})^2}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
  -2
  \frac{n(\mathbf{r'})}{4\pi
    R(\mathbf{r'})^3}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
\end{align}

Since the vector-weighted densities have no additional occurrences of
$R$, their derivatives come out similarly.
\begin{align}
  \mathbf{n}_{V2}(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))\frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} \\
  \frac{\delta \mathbf{n}_{V2}(\mathbf{r})}{\delta R(\mathbf{r}')}
  &= n(\mathbf{r'})\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|}
\end{align}
and analogously for $\mathbf{n}_{V1}$.
\begin{align}
  \mathbf{n}_{V1}(\mathbf{r}) &= \int \mathbf{dr'}
  \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}
  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))\frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} \\
  \frac{\delta \mathbf{n}_{V1}(\mathbf{r})}{\delta R(\mathbf{r}')}
  &= \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} -
  \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})^2}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|}
\end{align}

So, what is this convolution with a $\delta'$? Let's look at the
scalar version:
\begin{align}
  \bar{f}(\mathbf{r})
  &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
  \\
  &= \int
        f(\mathbf{r'})
        \lim_{\Delta R \rightarrow 0}
          \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
            -
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &= \lim_{\Delta R \rightarrow 0}
          \frac{\int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
            -
            \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &\approx\lim_{\Delta R \rightarrow 0}
          \frac{\int
            \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
            -
            \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &\approx\lim_{\Delta R \rightarrow 0}
          \frac{2\frac{\Delta R}{R} \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
      }{\Delta R}
  \\
  &= \frac{2}{R} \int
            f(\mathbf{r'})
            \delta(|\mathbf{r}-\mathbf{r'}| - R)
\end{align}
There is a sign ambiguity here with regard to whether we are asking
about the derivative of this delta function with regard to $R$ or
something else.  We just need to get it right.
